### 2015-11-23 ada-ftrl.tex
Added proofs for mirror descent section.

The regret bound can be optimized by choosing the optimal multiple of the
regularizer.  Namely, we choose regularizer of the form $\lambda f(w)$ where
$f(w)$ is $1$-strongly convex and optimize over $\lambda$. The result of the
optimization is the following corollary.  Its proof can be found in
Appendix~\ref{section:ada-ftrl-proof}.

@@

\begin{lemma}[Solution of the Recurrence]
\label{lemma:recurrence-solution}
Let $D$ be the diameter of $K$. The sequence $\{\Delta_t\}_{t=0}^\infty$
generated by AdaFTRL satisfies for any $T \ge 0$,
$$
\Delta_T \le \sqrt{3} \max\left\{D, \frac{1}{\sqrt{2\lambda}} \right\} \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \; .
$$
\end{lemma}
%
Proof of the Lemma~\ref{lemma:recurrence-solution} is deferred to
Appendix~\ref{section:ada-ftrl-proof}.
Theorem~\ref{theorem:ada-ftrl-regret-bound} follows from
Lemmas~\ref{lemma:initial-regret-bound}~and~\ref{lemma:recurrence-solution}.

>>>>

The regret bound can be optimized by choosing the optimal multiple of the
regularizer.  Namely, we choose regularizer of the form $\lambda f(w)$ where
$f(w)$ is $1$-strongly convex and optimize over $\lambda$. The result of the
optimization is the following corollary.  Its proof can be found in~\ref{section:ada-ftrl-proofs}.

@@

\begin{lemma}[Solution of the Recurrence]
\label{lemma:recurrence-solution}
Let $D$ be the diameter of $K$. The sequence $\{\Delta_t\}_{t=0}^\infty$
generated by AdaFTRL satisfies for any $T \ge 0$,
$$
\Delta_T \le \sqrt{3} \max\left\{D, \frac{1}{\sqrt{2\lambda}} \right\} \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} \; .
$$
\end{lemma}
%
Proof of the Lemma~\ref{lemma:recurrence-solution} is deferred to~\ref{section:ada-ftrl-proofs}.
Theorem~\ref{theorem:ada-ftrl-regret-bound} follows from
Lemmas~\ref{lemma:initial-regret-bound}~and~\ref{lemma:recurrence-solution}.

### 2015-11-23 mirror-descent.tex
Added proofs for mirror descent section.

\textsc{Mirror Descent} is a generic algorithm similar to \textsc{Follow The
Regularized Leader} but quite different in details. The algorithm is stated as
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
not differentiable\footnote{Note that this can happen even when $R_t$ is a
restriction of a differentiable function defined on a superset of $K$.  If $K$
is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
If $K$ is a subset of an affine subspace of a dimension smaller than the
dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, to define
Bregman divergence, $\Breg_{R_t}(u,v) = R(u) - R(v) - \langle \grad R(v), u - v
\rangle$, the algorithm has to choose together with $R_t$ a subgradient map
$\grad R_t:K \to V$.  If $R_t$ is a restriction of a differentiable function
$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
\in K$. The following lemma bounds regret of \textsc{Mirror Descent}.

\begin{lemma}[Regret of Mirror Descent]
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer} satisfies, for any
$u \in K$,
$$
\Regret_T(u)
\le
\sum_{t=1}^T \langle \ell_t, w_t - w_{t+1} \rangle - \Breg_{R_t}(w_{t+1}, w_t) + \Breg_{R_t}(u,w_t) - \Breg_{R_t}(u, w_{t+1}) \; .
$$
\end{lemma}

Several existing algorithms are special cases of \textsc{Mirror Descent}. Most
prominent example is projected gradient descent defined by $w_{t+1} = \Pi_K(w_t
- \eta_t \ell_t)$ where $\Pi_K$ is the Euclidean projection to $K$ and
$\eta_t>0$ is the step size.  Projected gradient descent is an instance of
\textsc{Mirror Descent} with regularizer $R_t(w) = \frac{1}{2\eta_t}\|w\|_2^2$.
A special case of projected gradient descent with step size $\eta_t =
\frac{1}{\sqrt{t}}$ is Zinkevich's \textsc{Generalized Infinitesimal Gradient
Ascent} (GIGA) algorithm~\cite{Zinkevich-2003}. Full-matrix and diagonal
versions of \textsc{AdaGrad with Composite Mirror Descent
Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of \textsc{Mirror
Descent} for any bounded $K$ with regularizers
$$
R_t(w) = \frac{1}{2} w^\top \left(\sum_{i=1}^t \ell_i \ell_i^\top \right)^{1/2} \!\!\!\! w
\qquad \text{and} \qquad
R_t(w) = \frac{1}{2} \sum_{j=1}^d w_j^2 \sqrt{ \sum_{i=1}^t \ell_{i,j}^2} \; .
$$
respectively. When $d=1$, both algorithms are equivalent to
\begin{equation}
\label{equation:ada-grad-regularizer}
R_t(w) = \frac{\|w\|_2^2}{2} \sqrt{\sum_{i=1}^t \|\ell_i\|_2^2 } \; .
\end{equation}
All three sequences of regularizers give a scale-free algorithm. Regularizer
\eqref{equation:ada-grad} can be viewed as a scale-free generalization of
Zinkevich's GIGA algorithm; in particular, when $\|\ell_i\|_2 = 1$ for all $i$,
\textsc{Mirror Descent} with regularizer \eqref{equation:ada-grad-regularizer}
coincides with GIGA.

\textsc{Exponentially-Weighted Average Forecaster}~\cite[Chapter
2]{Cesa-Bianchi-Lugosi-2006} with constant learning rate $\eta$ for the problem
of prediction with expert advice is an instance of \textsc{Mirror Descent} for
the probability simplex $K = \{ w \in \R^d ~:~ w_i \ge 0, \sum_{i=1}^d w_i = 0
\}$ with negative entropy regularizer $R_t(w) = \frac{1}{\eta} \sum_{i=1}^d w_i
\ln w_i$ where $R_t:K \to \R$. Note that this algorithm is \emph{not}
scale-free.

Replacing $\frac{1}{2}\|w\|_2^2$ in equation $\label{equation:ada-grad}$ with
an arbitrary convex function $R:K \to \R$ yields
\begin{equation}
\label{equation:scale-free-mirror-descent}
R_t(w) = R(w) \sqrt{\sum_{i=1}^t \|\ell_i\|_*^2} \; .
\end{equation}
We call the resulting algorithm \textsc{Scale-Free Mirror Descent} with
regularizer $R$.  Generalizing existing regret bound for
\textsc{AdaGrad}~\cite{Duchi-Hazan-Singer-2011} we can upper bound its regret.

@@

The second counter-example is as Theorem~\ref{theorem:second-counter-example}
below.  The decision set is the $d$-dimensional probability simplex $K = \{ w
\in \R^d ~:~ w_i \ge 0, \sum_{i=1}^d w_i = 0 \}$ and the regularizer is the
negative entropy $R(w) = \sum_{i=1}^d w_i \ln w_i$.  Negative entropy is
$1$-strongly convex with respect to $\|\cdot\|_1$ and the dual norm of
$\|\cdot\|_1$ is \mbox{$\|\cdot\|_\infty$}.  The corresponding Bregman
divergence is the Kullback-Leibler divergence $\Breg_R(u,v) = \sum_{i=1}^d u_i
\ln(u_i/v_i)$.  Note that despite that negative entropy is upper- and
lower-bounded, Kullback-Leibler divergence is not.

\begin{theorem}[Second Counter-Example]
\label{theorem:second-counter-example}
Let $d \ge 2$ and let $K = \{ w \in \R^d ~:~ w_i \ge 0, \sum_{i=1}^d w_i = 0 \}
\subseteq V$ be the $d$-dimensional probability simplex.  For any $T \ge 120$
there exists a sequence of loss vectors $\ell_1, \ell_2, \dots, \ell_T \in V^*$
such that $\|\ell_1\|_\infty = \|\ell_2\|_\infty = \dots = \|\ell_T\|_\infty =
1$ and \textsc{Scale-Free Mirror Descent} with regularizer $R(w) = \sum_{i=1}^d
w_i \ln w_i$ satisfies
$$
\Regret_T \ge \frac{T}{6} \; .
$$
\end{theorem}

>>>>

\textsc{Mirror Descent} is a generic algorithm similar to \textsc{Follow The
Regularized Leader} but quite different in details. The algorithm is stated as
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
not differentiable\footnote{Note that this can happen even when $R_t$ is a
restriction of a differentiable function defined on a superset of $K$.  If $K$
is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
If $K$ is a subset of an affine subspace of a dimension smaller than the
dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, Bregman
divergence, $\Breg_{R_t}(u,v) = R(u) - R(v) - \langle \grad R(v), u - v
\rangle$ needs to be defined. This is done by choosing a subgradient map $\grad
R_t:K \to V$, i.e. a function such that $\grad R_t(w)$ is a subgradient of
$R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
\in K$. The following lemma bounds regret of \textsc{Mirror Descent}.  Proof
can be found in~\ref{section:mirror-descent-proofs}.

\begin{lemma}[Regret of Mirror Descent]
\label{lemma:mirror-descent-regret}
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer} satisfies, for any
$u \in K$,
$$
\Regret_T(u)
\le
\sum_{t=1}^T \langle \ell_t, w_t - w_{t+1} \rangle - \Breg_{R_t}(w_{t+1}, w_t) + \Breg_{R_t}(u,w_t) - \Breg_{R_t}(u, w_{t+1}) \; .
$$
\end{lemma}

Several existing algorithms are special cases of \textsc{Mirror Descent}. Most
prominent example is projected gradient descent defined by $w_{t+1} = \Pi_K(w_t
- \eta_t \ell_t)$ where $\Pi_K$ is the Euclidean projection to $K$ and
$\eta_t>0$ is the step size.  Projected gradient descent is an instance of
\textsc{Mirror Descent} with regularizer $R_t(w) = \frac{1}{2\eta_t}\|w\|_2^2$.
A special case of projected gradient descent with step size $\eta_t =
\frac{1}{\sqrt{t}}$ is Zinkevich's \textsc{Generalized Infinitesimal Gradient
Ascent} (GIGA) algorithm~\cite{Zinkevich-2003}. Full-matrix and diagonal
versions of \textsc{AdaGrad with Composite Mirror Descent
Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of \textsc{Mirror
Descent} for any bounded $K$ with regularizers
$$
R_t(w) = \frac{1}{2} w^\top \left(\sum_{i=1}^t \ell_i \ell_i^\top \right)^{1/2} \!\!\!\! w
\qquad \text{and} \qquad
R_t(w) = \frac{1}{2} \sum_{j=1}^d w_j^2 \sqrt{ \sum_{i=1}^t \ell_{i,j}^2}
$$
respectively. When $d=1$, both algorithms are equivalent to
\begin{equation}
\label{equation:ada-grad-regularizer}
R_t(w) = \frac{\|w\|_2^2}{2} \sqrt{\sum_{i=1}^t \|\ell_i\|_2^2 } \; .
\end{equation}
All three sequences of regularizers give a scale-free algorithm. Regularizer
\eqref{equation:ada-grad-regularizer} can be viewed as a scale-free generalization of
Zinkevich's GIGA algorithm. In particular, when $\|\ell_i\|_2 = 1$ for all $i$,
\textsc{Mirror Descent} with regularizer \eqref{equation:ada-grad-regularizer}
coincides with GIGA.

\textsc{Exponentially-Weighted Average Forecaster}~\cite[Chapter
2]{Cesa-Bianchi-Lugosi-2006} with constant learning rate $\eta$ for the problem
of prediction with expert advice is an instance of \textsc{Mirror Descent} for
the probability simplex $K = \{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1
\}$ with negative entropy regularizer $R_t(w) = \frac{1}{\eta} \sum_{j=1}^d w_j
\ln w_j$ where $R_t:K \to \R$. Note that this algorithm is \emph{not}
scale-free.

Replacing $\frac{1}{2}\|w\|_2^2$ in equation \eqref{equation:ada-grad-regularizer} with
an arbitrary convex function $R:K \to \R$ yields
\begin{equation}
\label{equation:scale-free-mirror-descent}
R_t(w) = R(w) \sqrt{\sum_{i=1}^t \|\ell_i\|_*^2} \; .
\end{equation}
We call the resulting algorithm \textsc{Scale-Free Mirror Descent} with
regularizer $R$.  Generalizing existing regret bound for
\textsc{AdaGrad}~\cite{Duchi-Hazan-Singer-2011} we can upper bound its regret.
Proof can be found in~\ref{section:mirror-descent-proofs}.

@@

The second counter-example is stated as Theorem~\ref{theorem:second-counter-example}
below.  The decision set is the $d$-dimensional probability simplex $K = \{ w
\in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$ and the regularizer is the
negative entropy $R(w) = \sum_{j=1}^d w_j \ln w_j$.  Negative entropy is
$1$-strongly convex with respect to $\|\cdot\|_1$ and the dual norm of
$\|\cdot\|_1$ is \mbox{$\|\cdot\|_\infty$}.  The corresponding Bregman
divergence is the Kullback-Leibler divergence $\Breg_R(u,v) = \sum_{j=1}^d u_j
\ln(u_j/v_j)$.  Note that despite that negative entropy is upper- and
lower-bounded, Kullback-Leibler divergence is not.

\begin{theorem}[Second Counter-Example]
\label{theorem:second-counter-example}
Let $d \ge 2$ and let $K = \{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}
\subseteq V$ be the $d$-dimensional probability simplex.  For any $T \ge 120$
there exists a sequence of loss vectors $\ell_1, \ell_2, \dots, \ell_T \in V^*$
such that $\|\ell_1\|_\infty = \|\ell_2\|_\infty = \dots = \|\ell_T\|_\infty =
1$ and \textsc{Scale-Free Mirror Descent} with regularizer $R(w) = \sum_{j=1}^d
w_j \ln w_j$ satisfies
$$
\Regret_T \ge \frac{T}{6} \; .
$$
\end{theorem}

### 2015-11-23 per-coordinate.tex
Added proofs for mirror descent section.

Abstractly, we assume that the decision set is a Cartesian product $K=K_1
\times K_2 \times \dots \times K_d$ of a finite number of convex sets.  On each
factor $K_i$, $i=1,2,\dots,d$, we can run any OLO algorithm separately and we
denote by $\Regret_T^{(i)}(u_i)$ its regret with respect to $u_i \in K_i$. The
overall regret with respect to any $u=(u_1, u_2, \dots, u_d) \in K$ can be
written as
$$
\Regret_T(u) = \sum_{i=1}^d \Regret_T^{(i)}(u_i) \; .
$$
If the algorithm for each factor is scale-free, the overall algorithm is
clearly scale-free as well.  Using \textsc{AdaFTRL} or \textsc{SOLO FTRL} for
each factor $K_i$, we generalize and improve existing regret bounds
\cite{McMahan-Streeter-2010,Duchi-Hazan-Singer-2011} for algorithms with per-coordinate
learning rates.

>>>>

Abstractly, we assume that the decision set is a Cartesian product $K=K_1
\times K_2 \times \dots \times K_d$ of a finite number of convex sets.  On each
factor $K_j$, $i=1,2,\dots,d$, we can run any OLO algorithm separately and we
denote by $\Regret_T^{(j)}(u_j)$ its regret with respect to $u_j \in K_j$. The
overall regret with respect to any $u=(u_1, u_2, \dots, u_d) \in K$ can be
written as
$$
\Regret_T(u) = \sum_{j=1}^d \Regret_T^{(j)}(u_j) \; .
$$
If the algorithm for each factor is scale-free, the overall algorithm is
clearly scale-free as well.  Using \textsc{AdaFTRL} or \textsc{SOLO FTRL} for
each factor $K_j$, we generalize and improve existing regret bounds
\cite{McMahan-Streeter-2010,Duchi-Hazan-Singer-2011} for algorithms with per-coordinate
learning rates.

### 2015-11-23 preliminaries-proofs.tex
Added proofs for mirror descent section.

>>>>

\begin{proof}[Proof of Lemma~\ref{lemma:generic-regret-bound}]
By Fenchel-Young inequality,
\begin{align*}
\sum_{t=1}^T \left( R_{t+1}^*(-L_t) - R_t^*(-L_{t-1}) \right)
& = R_{T+1}^*(-L_T) - R_1^*(0) \\
& \ge - \langle L_T, u \rangle - R_{T+1}(u)  - R_1^*(0) \\
& = - R_{T+1}(u)  - R_1^*(0) - \sum_{t=1}^T \langle \ell_t, u \rangle \; .
\end{align*}
We add $\sum_{t=1}^T \langle \ell_t, w_t \rangle$ to both sides and we obtain
$\Regret_T(u)$ on the right side. After rearrangement of the terms, we get an upper bound
on the regret:
\begin{align*}
\Regret_T(u)
& = \sum_{t=1}^T \langle \ell_t, w_t \rangle - \sum_{t=1}^T \langle \ell_t, u \rangle \\
& \le R_{T+1}(u) + R_1^*(0) + \sum_{t=1}^T \left( R_{t+1}^*(-L_t) - R_t^*(-L_{t-1}) + \langle \ell_t, w_t \rangle \right) \; .
\end{align*}
By Proposition~\ref{proposition:conjugate-properties}, part 2, we have $w_t =
\grad R_t^*(-L_{t-1})$ and therefore we can rewrite the sum in last expression as
\begin{align*}
& \sum_{t=1}^T R_{t+1}^*(-L_t) - R_t^*(-L_{t-1}) + \langle \ell_t, w_t \rangle \\
& = \sum_{t=1}^T R_{t+1}^*(-L_t) - R_t^*(-L_{t-1}) + \langle \ell_t, \grad R_t^*(-L_{t-1}) \rangle \\
& = \sum_{t=1}^T R_t^*(-L_t) - R_t^*(-L_{t-1}) + \langle \ell_t, \grad R_t^*(-L_{t-1}) \rangle - R_t^*(-L_t) + R_{t+1}^*(-L_t) \\
& = \sum_{t=1}^T \Breg_{R_t^*}(-L_t, -L_{t-1}) - R_t^*(-L_t) + R_{t+1}^*(-L_t) \; .
\end{align*}
This finishes the proof.
\end{proof}

### 2015-11-23 preliminaries.tex
Added proofs for mirror descent section.

Except for properties 6 and 7, the proofs can be found
in~\cite{Shalev-Shwartz-2007}.  Property 6 is proven in
Appendix~\ref{section:definitions-proofs}. Property 7 trivially follows from
property 2.

@@

\begin{lemma}[Lemma 1 in \cite{Orabona-Crammer-Cesa-Bianchi-2014}]
\label{lemma:generic-regret-bound}
For any sequence $\{R_t\}_{t=1}^\infty$ of strongly convex lower
semi-continuous regularizers, regret of
Algorithm~\ref{algorithm:ftrl-varying-regularizer} is upper
bounded as
$$
\Regret_T(u) \le R_{T+1}(u) + R_1^*(0) + \sum_{t=1}^{T} \Breg_{R_t^*}(-L_t, -L_{t-1}) - R_t^*(-L_t) + R_{t+1}^*(-L_t) \; .
$$
\end{lemma}

\begin{proof}

\end{proof}

>>>>

Except for properties 6 and 7, the proofs can be found
in~\cite{Shalev-Shwartz-2007}.  Property 6 is proven
in~\ref{section:definitions-proofs}. Property 7 trivially follows from property
2.

@@

\begin{lemma}[Regret of Follow The Regularized Leader]
\label{lemma:generic-regret-bound}
For any sequence $\{R_t\}_{t=1}^\infty$ of strongly convex lower
semi-continuous regularizers, regret of
Algorithm~\ref{algorithm:ftrl-varying-regularizer} is upper
bounded as
$$
\Regret_T(u) \le R_{T+1}(u) + R_1^*(0) + \sum_{t=1}^{T} \Breg_{R_t^*}(-L_t, -L_{t-1}) - R_t^*(-L_t) + R_{t+1}^*(-L_t) \; .
$$
\end{lemma}

Proof of the lemma can be found in~\cite{Orabona-Crammer-Cesa-Bianchi-2014}. For completeness,
we include it in~\ref{section:definitions-proofs}.

### 2015-11-23 solo-ftrl-proofs.tex
Added proofs for mirror descent section.

\section{Proofs for SOLO FTRL}
\label{section:solo-ftrl-proof}

>>>>

\section{Proofs for SOLO FTRL}
\label{section:solo-ftrl-proofs}

\begin{proof}[Proof of Lemma~\ref{lemma:sum-of-square-roots-inverses}]
We use the inequality $x/\sqrt{x+y} \le 2(\sqrt{x+y} - \sqrt{y})$ which holds
for non-negative $x,y$ that are not both zero. Substituting $x = a_t$ and $y=\sum_{s=1}^{t-1} a_s$,
we get that for any $t \ge 1$,
$$
\frac{a_t}{\sqrt{\sum_{s=1}^t a_s}} \le 2 \sqrt{\sum_{s=1}^t a_s} \ - \ 2 \sqrt{\sum_{s=1}^{t-1} a_s} \; .
$$
Summing the above inequality over all $t=1,2,\dots,T$, the right side telescopes to
$2 \sqrt{\sum_{t=1}^T a_t}$.
\end{proof}

### 2015-11-23 solo-ftrl.tex
Few improvements
Added proofs for mirror descent section.

When $K$ is bounded, we can choose the optimal multiple of the regularizer.  We
choose $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex function and
optimize $\lambda$.  The result of the optimization is
Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}; the proof is in
Appendix~\ref{section:solo-ftrl-proof}.  It is similar to
Corollary~\ref{corollary:ada-ftrl-regret-bound} for AdaFTRL. The scaling
however is different in the two corollaries.  In
Corollary~\ref{corollary:ada-ftrl-regret-bound}, $\lambda \sim 1/(\sup_{v \in
K} f(v))$ while in Corollary~\ref{corollary:regret-solo-ftrl-bounded-set} we
have $\lambda \sim 1/\sqrt{\sup_{v \in K} f(v)}$.

\begin{corollary}[Regret Bound for Bounded Decision Sets]
\label{corollary:regret-solo-ftrl-bounded-set}
Suppose $K \subseteq V$ is a non-empty bounded closed convex subset.  Suppose
that $f:K \to \R$ is a non-negative lower semi-continuous function that is $1$-strongly
convex with respect to $\|\cdot\|$. SOLO FTRL with regularizer
$$
R(w) = \frac{f(w)\sqrt{2.75}}{\sqrt{\sup_{v \in K} f(v)}}
\quad \text{satisfies} \quad %
\Regret_T \le 13.3 \sqrt{\sup_{v \in K} f(v) \sum_{t=1}^{T} \norm{\ell_t}_*^2} \; .
$$
\end{corollary}

@@

\begin{lemma}[Useful Inequality]
\label{lemma:useful}
Let $C, a_1, a_2, \dots, a_T\geq0$. Then,
$$
\sum_{t=1}^T \min \left\{ a_t^2 / \sqrt{\sum_{s=1}^{t-1} a_s^2}, \ C a_t \right\}
\le 3.5 C \max_{t=1,2,\dots,T} a_t + 3.5 \sqrt{\sum_{t=1}^T a_t^2} \; .
$$
\end{lemma}
\begin{proof}
Without loss of generality, we can assume that $a_t > 0$ for all $t$. Since otherwise we
can remove all $a_t = 0$ without affecting either side of the inequality. Let $M_t = \max\{a_1, a_2, \dots, a_t\}$ and $M_0 = 0$.
We prove that for any $\alpha > 1$
$$
\min\left\{ \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}, C a_t \right\}
\le 2 \sqrt{1+\alpha^2} \left( \sqrt{\sum_{s=1}^t a_s^2} - \sqrt{\sum_{s=1}^{t-1} a_s^2} \right) + \frac{C\alpha( M_t  - M_{t-1})}{\alpha - 1}
$$
from which the inequality follows by summing over $t=1,2,\dots,T$ and choosing $\alpha = \sqrt{2}$.
The inequality follows by case analysis. If $a_t^2 \le \alpha^2 \sum_{s=1}^{t-1} a_s^2$, we have
\begin{multline*}
\min\left\{ \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}, C a_t \right\}
\le \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}
= \frac{a_t^2}{\sqrt{\frac{1}{1+\alpha^2} \left( \alpha^2 \sum_{s=1}^{t-1} a_s^2 + \sum_{s=1}^{t-1} a_s^2 \right)}} \\
\le \frac{a_t^2\sqrt{1+\alpha^2}}{\sqrt{ a_t^2 + \sum_{s=1}^{t-1} a_s^2 }}
= \frac{a_t^2\sqrt{1+\alpha^2}}{\sqrt{\sum_{s=1}^t a_s^2}}
\le 2\sqrt{1+\alpha^2} \left( \sqrt{\sum_{s=1}^t a_s^2} - \sqrt{\sum_{s=1}^{t-1} a_s^2} \right)
\end{multline*}
where we have used $x^2/\sqrt{x^2+y^2} \le 2(\sqrt{x^2+y^2} - \sqrt{y^2})$ in the last step.
On the other hand, if $a_t^2 > \alpha^2 \sum_{t=1}^{t-1} a_s^2$, we have
\begin{multline*}
\min\left\{ \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}, C a_t \right\}
\le C a_t
= C \frac{\alpha a_t  - a_t}{\alpha - 1}
\le \frac{C}{\alpha - 1} \left( \alpha a_t  - \alpha \sqrt{\sum_{s=1}^{t-1} a_s^2} \right) \\
= \frac{C\alpha}{\alpha - 1} \left( a_t  - \sqrt{\sum_{s=1}^{t-1} a_s^2} \right)
\le \frac{C\alpha}{\alpha - 1} \left( a_t  - M_{t-1} \right)
= \frac{C\alpha}{\alpha - 1} \left( M_t  - M_{t-1} \right)
\end{multline*}
where we have used that $a_t = M_t$ and $\sqrt{\sum_{s=1}^{t-1} a_s^2} \ge M_{t-1}$.
\end{proof}

\begin{lemma}[Lemma~3.5 in \cite{Auer-Cesa-Bianchi-Gentile-2002}]
\label{lemma:sum-of-square-roots-inverses}
Let $a_1, a_2, \dots, a_T$ be non-negative real numbers. If $a_1 > 0$ then,
$$
\sum_{t=1}^T a_t / \sqrt{\sum_{s=1}^t a_s} \le 2 \sqrt{\sum_{t=1}^T a_t} \; .
$$
\end{lemma}

>>>>

When $K$ is bounded, we can choose the optimal multiple of the regularizer.  We
choose $R(w) = \lambda f(w)$ where $f$ is a $1$-strongly convex function and
optimize $\lambda$.  The result of the optimization is
Corollary~\ref{corollary:regret-solo-ftrl-bounded-set}; the proof is in
Appendix~\ref{section:solo-ftrl-proofs}.  It is similar to
Corollary~\ref{corollary:ada-ftrl-regret-bound} for AdaFTRL. The scaling
however is different in the two corollaries.  In
Corollary~\ref{corollary:ada-ftrl-regret-bound}, $\lambda \sim 1/(\sup_{v \in
K} f(v))$ while in Corollary~\ref{corollary:regret-solo-ftrl-bounded-set} we
have $\lambda \sim 1/\sqrt{\sup_{v \in K} f(v)}$.

\begin{corollary}[Regret Bound for Bounded Decision Sets]
\label{corollary:regret-solo-ftrl-bounded-set}
Suppose $K \subseteq V$ is a non-empty bounded closed convex subset.  Suppose
that $f:K \to \R$ is a non-negative lower semi-continuous function that is $1$-strongly
convex with respect to $\|\cdot\|$. SOLO FTRL with regularizer
$$
R(w) = \frac{f(w)\sqrt{2.75}}{\sqrt{\sup_{v \in K} f(v)}}
\quad \text{satisfies} \quad
\Regret_T \le 13.3 \sqrt{\sup_{v \in K} f(v) \sum_{t=1}^{T} \norm{\ell_t}_*^2} \; .
$$
\end{corollary}

@@

\begin{lemma}[Useful Inequality]
\label{lemma:useful}
Let $C, a_1, a_2, \dots, a_T\geq0$. Then,
$$
\sum_{t=1}^T \min \left\{ \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}, \ C a_t \right\}
\le 3.5 C \max_{t=1,2,\dots,T} a_t + 3.5 \sqrt{\sum_{t=1}^T a_t^2} \; .
$$
\end{lemma}
\begin{proof}
Without loss of generality, we can assume that $a_t > 0$ for all $t$. Since otherwise we
can remove all $a_t = 0$ without affecting either side of the inequality. Let $M_t = \max\{a_1, a_2, \dots, a_t\}$ and $M_0 = 0$.
We prove that for any $\alpha > 1$
$$
\min\left\{ \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}, C a_t \right\}
\le 2 \sqrt{1+\alpha^2} \left( \sqrt{\sum_{s=1}^t a_s^2} - \sqrt{\sum_{s=1}^{t-1} a_s^2} \right) + \frac{C\alpha( M_t  - M_{t-1})}{\alpha - 1}
$$
from which the inequality follows by summing over $t=1,2,\dots,T$ and choosing $\alpha = \sqrt{2}$.
The inequality follows by case analysis. If $a_t^2 \le \alpha^2 \sum_{s=1}^{t-1} a_s^2$, we have
\begin{multline*}
\min\left\{ \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}, C a_t \right\}
\le \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}
= \frac{a_t^2}{\sqrt{\frac{1}{1+\alpha^2} \left( \alpha^2 \sum_{s=1}^{t-1} a_s^2 + \sum_{s=1}^{t-1} a_s^2 \right)}} \\
\le \frac{a_t^2\sqrt{1+\alpha^2}}{\sqrt{ a_t^2 + \sum_{s=1}^{t-1} a_s^2 }}
= \frac{a_t^2\sqrt{1+\alpha^2}}{\sqrt{\sum_{s=1}^t a_s^2}}
\le 2\sqrt{1+\alpha^2} \left( \sqrt{\sum_{s=1}^t a_s^2} - \sqrt{\sum_{s=1}^{t-1} a_s^2} \right)
\end{multline*}
where we have used $x^2/\sqrt{x^2+y^2} \le 2(\sqrt{x^2+y^2} - \sqrt{y^2})$ in the last step.
On the other hand, if $a_t^2 > \alpha^2 \sum_{t=1}^{t-1} a_s^2$, we have
\begin{multline*}
\min\left\{ \frac{a_t^2}{\sqrt{\sum_{s=1}^{t-1} a_s^2}}, C a_t \right\}
\le C a_t
= C \frac{\alpha a_t  - a_t}{\alpha - 1}
\le \frac{C}{\alpha - 1} \left( \alpha a_t  - \alpha \sqrt{\sum_{s=1}^{t-1} a_s^2} \right) \\
= \frac{C\alpha}{\alpha - 1} \left( a_t  - \sqrt{\sum_{s=1}^{t-1} a_s^2} \right)
\le \frac{C\alpha}{\alpha - 1} \left( a_t  - M_{t-1} \right)
= \frac{C\alpha}{\alpha - 1} \left( M_t  - M_{t-1} \right)
\end{multline*}
where we have used that $a_t = M_t$ and $\sqrt{\sum_{s=1}^{t-1} a_s^2} \ge M_{t-1}$.
\end{proof}

\begin{lemma}
\label{lemma:sum-of-square-roots-inverses}
Let $a_1, a_2, \dots, a_T$ be non-negative real numbers. If $a_1 > 0$ then,
$$
\sum_{t=1}^T \frac{a_t}{\sqrt{\sum_{s=1}^t a_s}} \le 2 \sqrt{\sum_{t=1}^T a_t} \; .
$$
\end{lemma}
%
Lemma~\ref{lemma:sum-of-square-roots-inverses} is from \cite[Lemma
3.5]{Auer-Cesa-Bianchi-Gentile-2002}.  We give
a proof in~\ref{section:solo-ftrl-proofs} for completeness.
